```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

This homework is going to focus on using R to make KNN predictions and somewhat reproduce the plots I used in lecture 7.

# Setup

For the lecture I used the following code to generate data:

```{r}
n = 50
set.seed(101)
x1 = runif(n)
x2 = runif(n)
prob  = ifelse(x1 < 0.5 & x1 > 0.25 & x2 > 0.25 & x2<0.75,0.8,0.3)
y  = as.factor(rbinom(n,1,prob))
levels(y) = c("1","2")
df = data.frame(y=y,x1=x1,x2=x2)
```

And then I used the following function to make KNN predictions:
```{r}
knn_pred = function(point,x1,x2,y,k) {
  dists = sqrt((x1-point[1])^2+(x2-point[2])^2) #Find all distances to current obs
  bound = sort(dists)[k]                #Find kth smallest distance
  indices = which(dists <= bound)       #Find which obs have dists 1:k
  outcomes = as.integer(y[indices])     #Find corresponding outcomes y
  round(mean(outcomes)) #Taking advantage of 2 outcomes. If more 2s, this gives 2, if more 1s this gives 1. 
}
```

This code builds a grid of points, and then makes predictions for each of those points.
```{r}
grid.fineness = 201
sequence = seq(0,1,length.out=grid.fineness)
grid = expand.grid(sequence,sequence)
colnames(grid) = c("x1","x2")
yhat = apply(grid,1,knn_pred,x1=x1,x2=x2,y=y,k=5)
yhat = as.factor(yhat)
```

With those predictions, we can build a dataframe, and plot.
```{r}
df = as.data.frame(grid)
df$y = yhat
ggplot(df,aes(x=x1,y=x2,col=y))+geom_point(size=0.08)
```

If we drop the `round` and subtract 1 in our `knn_pred` function, we can get probabilities out.
```{r}
knn_prob = function(point,x1,x2,y,k) {
  dists = sqrt((x1-point[1])^2+(x2-point[2])^2) #Find all distances to current obs
  bound = sort(dists)[k]                #Find kth smallest distance
  indices = which(dists <= bound)       #Find which obs have dists 1:k
  outcomes = as.integer(y[indices])     #Find corresponding outcomes y
  mean(outcomes)-1 #Taking advantage of 2 outcomes.
}
```

We can predict those probabilities at each point:
```{r}
phat = apply(grid,1,knn_prob,x1=x1,x2=x2,y=y,k=1)
df$phat = phat
ggplot(df,aes(x=x1,y=x2,col=phat))+geom_point(size=0.08)
```

This is, in essence, the beginnings of a simulation study. We generated data, and we can look at how our predictions perform. We can do this with either the classifications or the underlying probabilities. 

# Questions

We are going to extend this simulation study in a few ways. 

## Q1 - Bigger Sample

Resetting the seed with:
```{r}
set.seed(101)
```

Run the same data generation code, but with a sample size of 1000. Plot the resulting probabilities when we use $K=1$, $K=5$, and $K=25$. 

```{r}
n=1000
x1 = runif(n)
x2 = runif(n)
prob  = ifelse(x1 < 0.5 & x1 > 0.25 & x2 > 0.25 & x2<0.75, 0.8, 0.3)
y  = as.factor(rbinom(n,1,prob))
levels(y) = c("1","2")
df1 = data.frame(y=y,x1=x1,x2=x2)

grid.fineness = 201
sequence = seq(0,1,length.out=grid.fineness)
grid = expand.grid(sequence,sequence)
colnames(grid) = c("x1","x2")
yhat = apply(grid,1,knn_pred,x1=x1,x2=x2,y=y,k=1)
yhat = as.factor(yhat)

df = as.data.frame(grid)
df$y = yhat

phat = apply(grid,1,knn_prob,x1=x1,x2=x2,y=y,k=1)
df$phat = phat
ggplot(df,aes(x=x1,y=x2,col=phat))+geom_point(size=0.08)
```

```{r}
yhat = apply(grid,1,knn_pred,x1=x1,x2=x2,y=y,k=5)
yhat = as.factor(yhat)
df$y = yhat
phat = apply(grid,1,knn_prob,x1=x1,x2=x2,y=y,k=5)
df$phat = phat

ggplot(df,aes(x=x1,y=x2,col=phat))+geom_point(size=0.08)
```

```{r}
yhat = apply(grid,1,knn_pred,x1=x1,x2=x2,y=y,k=25)
yhat = as.factor(yhat)
df$y = yhat
phat = apply(grid,1,knn_prob,x1=x1,x2=x2,y=y,k=25)
df$phat = phat

ggplot(df,aes(x=x1,y=x2,col=phat))+geom_point(size=0.08)
```

Plot the classification *predictions* when $K=10$, using a probability threshold of 0.2 for our predictions instead of the standard 0.5. 

```{r, echo=T}
n=1000
x1 = runif(n)
x2 = runif(n)
prob  = ifelse(x1 < 0.5 & x1 > 0.25 & x2 > 0.25 & x2<0.75, 0.8, 0.3)
y  = as.factor(rbinom(n,1,prob))
levels(y) = c("1","2")
df = data.frame(y=y,x1=x1,x2=x2)

knn_pred = function(point,x1,x2,y,k) {
  dists = sqrt((x1-point[1])^2+(x2-point[2])^2) #Find all distances to current obs
  bound = sort(dists)[k]                #Find kth smallest distance
  indices = which(dists <= bound)       #Find which obs have dists 1:k
  outcomes = as.integer(y[indices])     #Find corresponding outcomes y
  round(mean(outcomes)) #Taking advantage of 2 outcomes. If more 2s, this gives 2, if more 1s this gives 1. 
}

grid.fineness = 201
sequence = seq(0,1,length.out=grid.fineness)
grid = expand.grid(sequence,sequence)
colnames(grid) = c("x1","x2")

phat = apply(grid,1,knn_prob,x1=x1,x2=x2,y=y,k=10)
yhat = as.factor((phat >= 0.2)+1)

df_grid = data.frame(x1=grid$x1,
                     x2=grid$x2,
                     y=yhat,
                     p=phat)
ggplot(df_grid,aes(x=x1,y=x2,col=y))+geom_point(size=0.08)
```
## Q2 -- Logit Comparison

Fit an interacted logit to this data. (i.e. model $Y\sim x1+x2+x1:x2$ -- using `glm`, not a LASSO). Find the predicted probabilities for every point in our grid, and plot those predicted probabilities. 

```{r}
set.seed(101)

n=1000
x1 = runif(n)
x2 = runif(n)
prob  = ifelse(x1 < 0.5 & x1 > 0.25 & x2 > 0.25 & x2<0.75, 0.8, 0.3)
y  = as.factor(rbinom(n,1,prob))
levels(y) = c("1","2")
df = data.frame(y=y,x1=x1,x2=x2)
dfp = df[c(-1)]

x = glm(y ~ x1 + x2 + x1:x2, data=df, family='binomial')

grid.fineness = 201
sequence = seq(0,1,length.out=grid.fineness)
grid = expand.grid(sequence,sequence)
colnames(grid) = c("x1","x2")
grid <- grid %>% 
  mutate(z = predict(x, newdata=grid, "response"))
ggplot(grid,aes(x=x1,y=x2,col=z))+geom_point(size=0.08)
```

## Q3 -- ROC

Plot the (in-sample) ROC curves for both the logit model and the KNN with $K=10$. (hint: I have a function for doing this, given outcomes and probabilities in the lecture)

Which of these models looks better? 

```{r}
roc = function(p,y, ...){
  y = factor(y)
  n = length(p)
  p = as.vector(p)
  probs = seq(0,1,length=1001)
  mat = matrix(rep(probs,n),ncol=length(probs),byrow=TRUE)
  Q = p > mat
  specificity = colMeans(!Q[y==levels(y)[1],])
  sensitivity = colMeans(Q[y==levels(y)[2],])
  plot(1-specificity, sensitivity, type="l", ...)
  abline(a=0,b=1,lty=2,col=8)
}

thresholds = seq(0,1,length.out=6)
Q = phat > matrix(rep(thresholds,n),ncol=length(thresholds),byrow=TRUE)

max.i = 1
specificity = colMeans(!Q[y==levels(y)[1],])[1:max.i]
sensitivity = colMeans(Q[y==levels(y)[2],])[1:max.i]
```
```{r}
max.i = 2
specificity = colMeans(!Q[y==levels(y)[1],])[1:max.i]
sensitivity = colMeans(Q[y==levels(y)[2],])[1:max.i]
plot(1-specificity,sensitivity,type="b",xlim=c(0,1),ylim=c(0,1),lwd=2)
abline(a=0,b=1,lty=2,col=8)
```


## Q4 -- Survey
Please complete the surveys posted last week (and in announcements on canvas). They will help me ensure the last 5 weeks of class are as useful to you as possible. 



# Optional

1. Plot the resulting predictions from a $k=30$ KNN for Q4.
2. Find OOS ROC curves in Q2 -- build a holdout, then use predictions on it. 3. Do K-fold cross validation for the OOS ROC curves
4. Add squared terms to the logit, remake predictions and plots (e.g. `y~x1+x1^2+x2+x2^2+x1:x2)`

## Optional Long Q

We have a new output with three categories.
```{r}
# x1,x2,y from your sample with 1k observations need to exist to run this.
set.seed(101)
z = ifelse((x1>0.8 | x2 < 0.4),rbinom(length(y),1,0.8),(y==1)*2)
```
Modify `knn_pred` so that it predicts the most likely category out of 3 categories (hint: the functions `table`, `which.max`, `names`, and `as.integer` are how I did this) (hint2: maybe start by building a function that takes a vector and finds the most common element, then fit it into the rest of this). Plot the grid-predictions with this new classifier.





